### 2023-11-26 report/paper.tex
add initial references

\documentclass[12pt]{article}
\usepackage[legalpaper, portrait, margin=1in]{geometry}
\usepackage{multicol}
\usepackage[english]{babel}

\usepackage{xcolor}
\usepackage{blindtext}

\title{Automatic Lyric Transcription}
\author{Victor Shan}
\begin{document}
\color{red}
\maketitle
\begin{multicols*}{2}
\begin{abstract}
\blindtext
\end{abstract}

\section*{Scholarly Articles}
\begin{enumerate}
    \item \cite{wav2vecTransfer}
    \item \cite{wav2vec}
    \item \cite{Librispeech}
    \item \cite{whisper}
    \item \cite{LyricAlignAcoustic}
    \item \cite{Audio2Score}
    \item \cite{PhonemeInformed}
    \item \cite{Hansen}
    \item \cite{KaraokeApp}
    \item \cite{BayesianAST}
    \item \cite{whisperX}
    \item \cite{CSD}
    \item \cite{DALI}
    \item \cite{NUSDataset}
    \item \cite{JamendoLyrics}
    \item \cite{musDB18}
    \item \cite{Demux}
    \item \cite{CMUDict}
    \item \cite{SpecAugment}
    \item \cite{PhonationMode}
\end{enumerate}
\section*{Other Sources}
\begin{enumerate}
    \item \cite{MFCC}
    \item \cite{CTC}
\end{enumerate}
\bibliographystyle{apalike}
\bibliography{references}
\end{multicols*}
\end{document}

### 2023-11-27 report/paper.tex
add references and basic outline
fill out dataset sections

@@
 \usepackage[english]{babel}
+\usepackage{graphicx}
 
@@
 
-\title{Automatic Lyric Transcription}
+% Paper requirement 1.5 line spacing. Remove for github publishing
+\usepackage{setspace}
+\onehalfspacing
+
+\title{Automatic Singing Transcription}
 \author{Victor Shan}
 \begin{document}
-\color{red}
 \maketitle
@@
 \begin{abstract}
-\blindtext
 \end{abstract}
 
-\section*{Scholarly Articles}
+\section{Introduction}
+
+Songs play an important part of society because they are conduits of cultural expression,
+foster emotional connection and is a great source of entertainment. Automatic Singing Transcription
+(AST) is a branch of Automatic Speech Recognition (ASR) can facilitate a deeper way to interact with
+it. In ASR, the goal is to transcribe normal speech to text,
+whether that's a conversation between two people, a lecture, or a podcast. Examples of ASR include
+Youtube's auto-generated captions, assistants like Siri and Alexa, and even live captioning of
+meetings through applications like Zoom. AST is a similar task, but instead of transcribing normal
+speech, AST transcribes singing. Applications of AST range from karaoke applications to Music
+Information Retrieval (MIR) systems. These MIR systems can be used to search for songs based on
+lyrics, categorize songs based on their lyrics, or even aid in the generation of new songs.
+
+While AST is similar to ASR, there are a few key differences that make it a more challenging task.
+First, singing is a more complex signal than normal speech. Singing has a more dynamic signal, with
+more variation in pitch, duration and vibrato \cite{SemiSupervisedSolo}.
+For example, vowels sounds are often held for longer in singing than in normal speech like dragging
+the last part of ``bye" in word ``goodbye". Songs also often include music in the background, which can
+make it more difficult to distinguish between the singer's voice and the background music. Finally,
+there is a lack of large, publicly available datasets for AST \cite{DALI}. This is in contrast to ASR, where
+there are many large datasets available, such as LibriSpeech \cite{Librispeech}. This lack of data
+makes it especially difficult to train AST models.
+
+There are many possible outputs for AST but one of the most detailed and useful outputs is the
+time-aligned phoneme sequence. This is a sequence of phonemes, or the smallest unit of sound in a
+language, that is aligned with the audio. Time-aligned phoneme sequences are useful because it can contains enough
+information to reconstruct the not only the lyrics, but also how they match with the music. This
+output can be used to generate a karaoke application that highlights the lyrics as they are sung,
+or even train a model to sing covers of songs. Word-level alignement is an easier alternative but it
+misses the crucial information of the duration of phonemes within words. The ideal output is one that includes even more information
+such as the notes to sing at so that singers would be able to read and sing the song without any
+additional information. In this paper, we will mainly focus on the phoneme-level alignment but use
+word-level alignment to compare the performance of our model to other models.
+
+\section{Datasets}
+
+\subsection{Requirements}
+The requirements of a good dataset are audio files, lyrics and timestamps. The audio files can be
+in any format such as mp3 or wav. The lyrics can be in any format as well, but the timestamps must
+be in a format that can be used to align the lyrics with the audio. The timestamps can be in the
+form of a phoneme sequence or word sequence but should also include the onset and offeset times
+(beginning and end). Sentence level timestamps are too coarse and would need
+to be broken down into word or phoneme level timestamps. Most lyrics used by popular music services
+such as Spotify only use sentence/line level timestamps. Ideally, the dataset would also be large
+enough to train deep neural networks and publicly available so that other researchers can use it and
+compare their models to existing models.
+
+\subsection{Challenges}
+With these requirements in mind, it is easy to see why there are so few datasets available. First,
+it is difficult to create a dataset with lyrics and timestamps. The lyrics must be manually
+transcribed and the timestamps must be manually aligned with the audio. This is a time consuming
+process that requires a lot of effort. Second, it is difficult to obtain the rights to use the
+audio files. Most songs are owned by record labels and it is difficult to obtain the rights to use
+these files. These two challenges make it difficult to create new datasets and is the reason why
+there are so few datasets available and why most of the existing datasets are small.
+
+\subsection{Dataset Augmentation}
+\subsubsection{SpecAugment}
+Due to the lack of datasets, it is more important to make the most of the existing datasets. One
+way to do this is to augment the existing datasets. SpecAugment is a series of techniques that
+augment the audio spectrogram to improve the performance of ASR models. A frequent intermediate
+representation are Mel-Frequency Cepstral Coefficients (MFCC). These are image representations of
+energy at certain frequencies on a scale that more closely matches human hearing \cite{MFCC}.
+These techniques include time warping, frequency masking and time masking \cite{SpecAugment}.
+Time warping is a technique that stretches or compresses the audio spectrogram in the time
+dimension. Frequency masking is a technique that masks a random number of frequency channels in
+the audio spectrogram. Time masking is a technique that masks a random number of time steps
+in the audio spectrogram. These techniques are used to augment the audio spectrogram before it
+is fed into the ASR model. \cite{SpecAugment}
+
+% ----------------- INCLUDE MFCC IMAGE -----------------
+
+\subsubsection{Transforming Existing Datasets Into Pseudo-Singing Datasets} \label{sec:pseudoSinging}
+Another way to augment the existing datasets is to transform the existing datasets. This can be
+done by shifting the pitch, duration or vibrato \cite{SongifiedSpeech}. The advantage of this
+technique is that is can also be applied to speech datasets and transform them into pseudo-singing
+datasets. The disadvantage is that the results will contain artifacts from the transformation based
+on the techiniques applied. Neural network models showed an almost 15\% improvement on the
+transformed TIMIT dataset than ones trained on the original TIMIT dataset \cite{SongifiedSpeech}.
+
+\subsubsection{Transforming Utterance level datasets into Phoneme level datasets} \label{sec:utteranceToPhoneme}
+A technique from ASR that can also be applied to AST is transforming utterance/sentence
+level datasets into phoneme level datasets. There are many ASR datasets that contain single utterances
+such as LibriSpeech and MUSDB18 \cite{Librispeech,musdb18}. These datasets can be effectively
+transformed into phoneme level datasets by using a phoneme dictionary such as the CMU Pronouncing
+Dictionary \cite{CMUDict}. This dictionary contains a mapping from words to phonemes. Using this
+dictionary, the utterances can be transformed into phoneme sequences. These sequences can then be
+used with a Connectionist Temporal Classification (CTC) loss function to train AST models \cite{CTC}.
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.5\textwidth]{assets/CTC.png}
+    \caption{CTC Function collapsing repetitions \cite{CTC}}
+    \label{fig:CTC}
+\end{figure*}
+CTC allows for the model to output a sequence of phonemes per time step and the
+repetitions are collapsed into a single phoneme as shown in Figure \ref{fig:CTC}.
+This technique was used to allow a model to incorperate the utterance level LibriSpeech
+dataset into the training for a phoneme ASR model \cite{wav2vec}. Timestamps can be
+retrieved from the pre-CTC output that had time-aligned phoneme classifications.
+The same process can be applied to AST models to allow them to use the utterance level song datasets.
+
+\subsubsection{Teacher-Student Approach} \label{sec:TeacherStudent}
+The teacher-student approach is inspired by the technique of the same name that was intended
+to reduce the size of large deep neural networks. The idea was first train a large deep neural
+network, the teacher, and then train a smaller deep neural network, the student, to mimic the
+teacher \cite{TeacherStudent}. However, this technique can also be applied in cases of low labeled
+data availablility but high unlabeled data availablility. To start off, a model would be
+trained on a small dataset of labeled data. Then, the model would be used to label a large dataset
+of unlabeled data. Finally, a new model would be trained on the newly labeled dataset
+\cite{DALI}. This has proven to be effective for transcribing drums in music with the student
+model outperforming the teacher model \cite{DrumsStudentTeacher}.
+
+\subsection{Existing Datasets}
+
+\subsubsection{TIMIT (ASR)}
+TIMIT is a dataset of speech recordings of 630 speakers of eight major dialects of American English
+with time-aligned phoneme sequences \cite{TIMIT}. It is a popular dataset for ASR and has been used
+to train many ASR models. However, since this is not a singing dataset, it does not contain any of
+the characteristics of singing such as pitch, duration or vibrato. This is an excellent dataset to
+apply transformation into a pseudo-singing dataset metioned in \ref{sec:pseudoSinging}. The popularity of this dataset
+makes it a good candidate for applying transformations to create a pseudo-singing dataset and also
+for a general benchmark to compare against other models.
+
+\subsubsection{LibriSpeech (ASR)}
+LibriSpeech is a 1000 hour dataset of audiobook recordings where each recording has a matching sentence \cite{Librispeech}. This
+dataset is also a popular dataset for ASR and has been used to train many ASR models including
+recent breakthroughs like wav2vec 2.0 \cite{wav2vec}. This dataset is also a good candidate for
+training AST models because it is large and publicly available. This dataset is special because of
+the clarity of the audio and previous success by other models such as the wav2vec 2.0 model in
+detecting phonemes in this dataset when fine tuned with the TIMIT dataset \cite{wav2vec}. This is
+done through the CTC technique mentioned in \ref{sec:utteranceToPhoneme}. Since
+speech and singing both use the same phonemes, because they are in the same language, this large
+dataset can be used to train base models before fine-tuning them AST models.
+
+\subsubsection{Jamendo Dataset}
+This dataset one of the most popular datasets for AST and has been used by many state-of-the-art
+AST models. It contains 20 English songs and 60 songs in other languages with word-aligned timestamped sequences \cite{JamendoLyrics}. This dataset
+is a good candidate for training AST models because it is publicly available and it's popularity
+makes it an excellent benchmark to compare against other models. However, it is still a relatively
+small dataset and does not contain any phoneme sequences. The authors of this dataset were able to
+achieve a 77.8\% Word Error Rate (WER) which still leaves a lot of room for improvement.
+{
+\color{red}
+Remember to record the performance of the state of the art models on this dataset
+}
+% \subsubsection{Children's Songs Dataset} All Korean :(
+\subsubsection{MUSDB18}
+MUSEDB18 is a dataset of 150 songs with isolated vocals and accompaniment tracks \cite{musdb18}. This
+dataset is a good candidate for training AST models because it is publicly available and it has
+clean isolated vocals. The downside of this dataset is that it doesn't contain any word level
+timestamps. However, with the CTC technique mentioned in \ref{sec:utteranceToPhoneme}, this dataset
+can still be used to train AST models.
+
+\subsubsection{NUS Dataset}
+This dataset is one of the few datasets that contains phoneme level timestamps \cite{NUSDataset}.
+There are 169 minutes of 20 unique English songs by 12 different people. The CMUDict was used
+for the phoneme vocabulary and timestamps were manually annotated. This dataset is the ideal dataset
+type for training AST models due to this level of detail. It also includes a mix of slow to fast
+melodies annd a balanced gender distribution \cite{NUSDataset}.
+
+\subsubsection{Free Music Archive}
+The Free Music Archive (FMA) dataset is a dataset of 106,574 tracks with 161 genres \cite{FMA}. This
+dataset is not a good candidate for direclty training AST models because it does not contain any lyrics
+and some music may not even contain any singing at all. However, it is a good source of unlabeled
+songs that could be labeled through the teacher-student technique in \ref{sec:TeacherStudent}. It
+can also provide a good source of general singing audio that can be used in the
+training of wav2vec2.0 models that will be described in section \ref{sec:wav2vec}.
+
+\subsubsection{VocalSet}
+VocalSet is a 10 hour dataset of a capella singing from 20 professional singers
+demonstrating a variety of singing techniques \cite{VocalSet}. This dataset is a good candidate for
+training Voice Activity Detection (VAD) models because it contains onset and offset timestamps for
+each vocal segment. This is also a good a good dataset to help train AST models to know what singing
+sounds like.
+
+\subsubsection{Other Datasets}
+Many datasets were considered but left out due to the lack of availablility. Some of the most
+popular datasets such as Mauch's Dataset \cite{mirex2021} and Hansen's Dataset \cite{Hansen} are
+not publicly available anymore. Newer datasets such as DALI \cite{DALI} and DAMP! \cite{DAMP} are
+are hidden behind institutional logins and require manually requesting access.
+
+While both the Mauch's Dataset and Hansen's Datasets are quite small (Mauch has 20 songs, Hansen has 9 \cite{mirex2021}), the
+DALI and DAMP! datasets are much larger. The DALI dataset in particular used a version of the
+Teacher Student technique mentioned in \ref{sec:TeacherStudent} to label 105 songs with timestamps
+for the word and phoneme level \cite{DALI}. The DAMP! dataset is even larger with 300x30x2 song
+dataset. Both of these datasets would be excellent candidates for training AST models from their
+size alone.
+
+\section{Related Works}
+\subsection{Early HMM-Based Acoustic Models}
+
+\subsection{Music Informed Models}
+\subsection{wav2vec 2.0 and Transfer Learning} \label{sec:wav2vec}
+\subsection{Whisper Word-Level Alignment}
+
+\section{Method}
 \begin{enumerate}
-    \item \cite{wav2vecTransfer}
-    \item \cite{wav2vec}
-    \item \cite{Librispeech}
-    \item \cite{whisper}
-    \item \cite{LyricAlignAcoustic}
-    \item \cite{Audio2Score}
-    \item \cite{PhonemeInformed}
-    \item \cite{Hansen}
-    \item \cite{KaraokeApp}
-    \item \cite{BayesianAST}
-    \item \cite{whisperX}
-    \item \cite{CSD}
-    \item \cite{DALI}
-    \item \cite{NUSDataset}
-    \item \cite{JamendoLyrics}
-    \item \cite{musDB18}
-    \item \cite{Demux}
-    \item \cite{CMUDict}
-    \item \cite{SpecAugment}
-    \item \cite{PhonationMode}
+    \item Preprocess datasets
+    \item Augment datasets
+    \item Create frankenstein dataset
+    \item Train model
+    \item Evaluate model
+    \item Label unlabeled datasets
+    \item train student model on newly labeled datasets
+    \item evaluate student model on original, manually labeled datasets
+    \item repeat
 \end{enumerate}
-\section*{Other Sources}
-\begin{enumerate}
-    \item \cite{MFCC}
-    \item \cite{CTC}
-\end{enumerate}
+
+\section{Results}
+
+\subsection{WER}
+
+\subsection{PER}
+
+\section{Discussion}
+
+\section{Future Work}
+\subsection{Adversarial Training}
+
+\section{Conclusion}
+
+% \section*{Scholarly Articles}
+% \begin{enumerate}
+%     \item \cite{wav2vecTransfer}
+%     \item \cite{wav2vec}
+%     \item \cite{Librispeech}
+%     \item \cite{whisper}
+%     \item \cite{LyricAlignAcoustic}
+%     \item \cite{Audio2Score}
+%     \item \cite{PhonemeInformed}
+%     \item \cite{Hansen}
+%     \item \cite{KaraokeApp}
+%     \item \cite{BayesianAST}
+%     \item \cite{whisperX}
+%     \item \cite{CSD}
+%     \item \cite{DALI}
+%     \item \cite{NUSDataset}
+%     \item \cite{JamendoLyrics}
+%     \item \cite{musDB18}
+%     \item \cite{Demux}
+%     \item \cite{CMUDict}
+%     \item \cite{SpecAugment}
+%     \item \cite{PhonationMode}
+% \end{enumerate}
+% \section*{Other Sources}
+% \begin{enumerate}
+%     \item \cite{MFCC}
+%     \item \cite{CTC}
+% \end{enumerate}
 \bibliographystyle{apalike}

### 2023-12-01 report/paper.tex
add to wav2vec 2

@@
-\documentclass[12pt]{article}
-\usepackage[legalpaper, portrait, margin=1in]{geometry}
+\documentclass[letterpaper, 12pt]{article}
+\usepackage[letterpaper, portrait, margin=1in]{geometry}
 \usepackage{multicol}
@@
 \section{Related Works}
-\subsection{Early HMM-Based Acoustic Models}
+\subsection{HMM Based Acoustic Models}
+The traditional approach to AST is can be separated into a pipeline of distinct steps. The first
+step is to extract the features from the audio. This usually includes some form of spectral analysis
+such as MFCCs. The second step is to use an acoustic model to classify the features into phonemes
+and generate a sequence of phonemes using a Hidden Markov Model (HMM). This is the traditional
+approach to ASR and is also used in AST \cite{Hansen}.
 
-\subsection{Music Informed Models}
+{
+\color{red}
+TODO: Add more detail about HMMs
+TODO: Write about Dynamic Time Warping
+TODO: Write about some results of HMMs
+}
+
+% \subsection{Music Informed Models}
+
 \subsection{wav2vec 2.0 and Transfer Learning} \label{sec:wav2vec}
+
+\subsubsection{wav2vec 2.0}
+One of the most recent breakthroughs in ASR is wav2vec 2.0 \cite{wav2vec}. This model uses a
+self-supervised learning approach was used during the training of the model. Unlabeled audio was
+fed into the model to learn discrete speech units. These discrete speech units required the
+Gumbel-Softmax \cite{gumbelSoftmax} to allow for backpropagation. The model was then
+fine-tuned with a linear layer and CTC loss on labeled data to perform ASR \cite{wav2vec}.
+
+This approach was able to achieve state-of-the-art results on the LibriSpeech dataset \cite{Librispeech}
+and the TIMIT dataset \cite{TIMIT}. However, the thing that makes this approach the most promising
+is the fact that after pre-training on a large amount of unlabeled data, the model can be fine-tuned
+on a 10 minute subset of labeled data to achieve 5.2 WER on the LibriSpeech clean dataset \cite{wav2vec}.
+This is very important for AST because there are so few labeled datasets available.
+
+\subsubsection{wav2vec 2.0 Transfer Learning}
+Using transfer learning for AST using wav2vec 2.0 was attempted in 2022 and was used to achieve
+state-of-the-art results on the Jamendo dataset as well as on the DALI, Hansen, Mauch,
+and DAMP! datasets \cite{wav2vecTransfer}. The approach they performed did not change the first part
+of the model that generated the discrete speech units. Instead, they changed the last part of the
+model to have another branch that outputs the probability of the current word given previous words
+and context. This approach achieved a 33.13 WER on the Jamendo dataset \cite{wav2vecTransfer}.
+
+Apart from their results, there were also a few other interesting things about their approach. Since
+the initial wav2vec 2.0 model was trained on the LibriSpeech dataset, they wanted to make the input
+audio of their model as similar to the LibriSpeech dataset as possible. To do this, they used
+Demucs v3 \cite{Demux} to separate the vocals from the accompaniment. They did not futher remove any
+noise or singing specific features from the audio. For their labeled datasets, they used
+utterance level labels excluding the instrumental parts. For the output, they used character level
+labels.
+
 \subsection{Whisper Word-Level Alignment}
@@
 \subsection{Adversarial Training}
+\subsection{Transform Singing to Speech Data}
+remove pitch from singing
 
@@
 
-% \section*{Scholarly Articles}
-% \begin{enumerate}
-%     \item \cite{wav2vecTransfer}
-%     \item \cite{wav2vec}
-%     \item \cite{Librispeech}
-%     \item \cite{whisper}
-%     \item \cite{LyricAlignAcoustic}
-%     \item \cite{Audio2Score}
-%     \item \cite{PhonemeInformed}
-%     \item \cite{Hansen}
-%     \item \cite{KaraokeApp}
-%     \item \cite{BayesianAST}
-%     \item \cite{whisperX}
-%     \item \cite{CSD}
-%     \item \cite{DALI}
-%     \item \cite{NUSDataset}
-%     \item \cite{JamendoLyrics}
-%     \item \cite{musDB18}
-%     \item \cite{Demux}
-%     \item \cite{CMUDict}
-%     \item \cite{SpecAugment}
-%     \item \cite{PhonationMode}
-% \end{enumerate}
-% \section*{Other Sources}
-% \begin{enumerate}
-%     \item \cite{MFCC}
-%     \item \cite{CTC}
-% \end{enumerate}
 \bibliographystyle{apalike}

### 2023-12-01 report/paper.tex
update to biblatex

@@
 \usepackage{blindtext}
+
+\usepackage[%
+backend=biber,
+natbib=true,
+% style=apa,
+style=ieee
+]{biblatex}
+\addbibresource{references.bib}
 
@@
 First, singing is a more complex signal than normal speech. Singing has a more dynamic signal, with
-more variation in pitch, duration and vibrato \cite{SemiSupervisedSolo}.
+more variation in pitch, duration and vibrato \citep{SemiSupervisedSolo}.
 For example, vowels sounds are often held for longer in singing than in normal speech like dragging
@@
 make it more difficult to distinguish between the singer's voice and the background music. Finally,
-there is a lack of large, publicly available datasets for AST \cite{DALI}. This is in contrast to ASR, where
-there are many large datasets available, such as LibriSpeech \cite{Librispeech}. This lack of data
+there is a lack of large, publicly available datasets for AST \citep{DALI}. This is in contrast to ASR, where
+there are many large datasets available, such as LibriSpeech \citep{Librispeech}. This lack of data
 makes it especially difficult to train AST models.
@@
 representation are Mel-Frequency Cepstral Coefficients (MFCC). These are image representations of
-energy at certain frequencies on a scale that more closely matches human hearing \cite{MFCC}.
-These techniques include time warping, frequency masking and time masking \cite{SpecAugment}.
+energy at certain frequencies on a scale that more closely matches human hearing \citep{MFCC}.
+These techniques include time warping, frequency masking and time masking \citep{SpecAugment}.
 Time warping is a technique that stretches or compresses the audio spectrogram in the time
@@
 in the audio spectrogram. These techniques are used to augment the audio spectrogram before it
-is fed into the ASR model. \cite{SpecAugment}
+is fed into the ASR model. \citep{SpecAugment}
 
@@
 Another way to augment the existing datasets is to transform the existing datasets. This can be
-done by shifting the pitch, duration or vibrato \cite{SongifiedSpeech}. The advantage of this
+done by shifting the pitch, duration or vibrato \citep{SongifiedSpeech}. The advantage of this
 technique is that is can also be applied to speech datasets and transform them into pseudo-singing
@@
 on the techiniques applied. Neural network models showed an almost 15\% improvement on the
-transformed TIMIT dataset than ones trained on the original TIMIT dataset \cite{SongifiedSpeech}.
+transformed TIMIT dataset than ones trained on the original TIMIT dataset \citep{SongifiedSpeech}.
 
@@
 level datasets into phoneme level datasets. There are many ASR datasets that contain single utterances
-such as LibriSpeech and MUSDB18 \cite{Librispeech,musdb18}. These datasets can be effectively
+such as LibriSpeech and MUSDB18 \citep{Librispeech,musdb18}. These datasets can be effectively
 transformed into phoneme level datasets by using a phoneme dictionary such as the CMU Pronouncing
-Dictionary \cite{CMUDict}. This dictionary contains a mapping from words to phonemes. Using this
+Dictionary \citep{CMUDict}. This dictionary contains a mapping from words to phonemes. Using this
 dictionary, the utterances can be transformed into phoneme sequences. These sequences can then be
-used with a Connectionist Temporal Classification (CTC) loss function to train AST models \cite{CTC}.
+used with a Connectionist Temporal Classification (CTC) loss function to train AST models \citep{CTC}.
 \begin{figure*}
@@
     \includegraphics[width=0.5\textwidth]{assets/CTC.png}
-    \caption{CTC Function collapsing repetitions \cite{CTC}}
+    \caption{CTC Function collapsing repetitions \citep{CTC}}
     \label{fig:CTC}
@@
 This technique was used to allow a model to incorperate the utterance level LibriSpeech
-dataset into the training for a phoneme ASR model \cite{wav2vec}. Timestamps can be
+dataset into the training for a phoneme ASR model \citep{wav2vec}. Timestamps can be
 retrieved from the pre-CTC output that had time-aligned phoneme classifications.
@@
 network, the teacher, and then train a smaller deep neural network, the student, to mimic the
-teacher \cite{TeacherStudent}. However, this technique can also be applied in cases of low labeled
+teacher \citep{TeacherStudent}. However, this technique can also be applied in cases of low labeled
 data availablility but high unlabeled data availablility. To start off, a model would be
@@
 of unlabeled data. Finally, a new model would be trained on the newly labeled dataset
-\cite{DALI}. This has proven to be effective for transcribing drums in music with the student
-model outperforming the teacher model \cite{DrumsStudentTeacher}.
+\citep{DALI}. This has proven to be effective for transcribing drums in music with the student
+model outperforming the teacher model \citep{DrumsStudentTeacher}.
 
@@
 TIMIT is a dataset of speech recordings of 630 speakers of eight major dialects of American English
-with time-aligned phoneme sequences \cite{TIMIT}. It is a popular dataset for ASR and has been used
+with time-aligned phoneme sequences \citep{TIMIT}. It is a popular dataset for ASR and has been used
 to train many ASR models. However, since this is not a singing dataset, it does not contain any of
@@
 \subsubsection{LibriSpeech (ASR)}
-LibriSpeech is a 1000 hour dataset of audiobook recordings where each recording has a matching sentence \cite{Librispeech}. This
+LibriSpeech is a 1000 hour dataset of audiobook recordings where each recording has a matching sentence \citep{Librispeech}. This
 dataset is also a popular dataset for ASR and has been used to train many ASR models including
-recent breakthroughs like wav2vec 2.0 \cite{wav2vec}. This dataset is also a good candidate for
+recent breakthroughs like wav2vec 2.0 \citep{wav2vec}. This dataset is also a good candidate for
 training AST models because it is large and publicly available. This dataset is special because of
 the clarity of the audio and previous success by other models such as the wav2vec 2.0 model in
-detecting phonemes in this dataset when fine tuned with the TIMIT dataset \cite{wav2vec}. This is
+detecting phonemes in this dataset when fine tuned with the TIMIT dataset \citep{wav2vec}. This is
 done through the CTC technique mentioned in \ref{sec:utteranceToPhoneme}. Since
@@
 This dataset one of the most popular datasets for AST and has been used by many state-of-the-art
-AST models. It contains 20 English songs and 60 songs in other languages with word-aligned timestamped sequences \cite{JamendoLyrics}. This dataset
+AST models. It contains 20 English songs and 60 songs in other languages with word-aligned timestamped sequences \citep{JamendoLyrics}. This dataset
 is a good candidate for training AST models because it is publicly available and it's popularity
@@
 \subsubsection{MUSDB18}
-MUSEDB18 is a dataset of 150 songs with isolated vocals and accompaniment tracks \cite{musdb18}. This
+MUSEDB18 is a dataset of 150 songs with isolated vocals and accompaniment tracks \citep{musdb18}. This
 dataset is a good candidate for training AST models because it is publicly available and it has
@@
 \subsubsection{NUS Dataset}
-This dataset is one of the few datasets that contains phoneme level timestamps \cite{NUSDataset}.
+This dataset is one of the few datasets that contains phoneme level timestamps \citep{NUSDataset}.
 There are 169 minutes of 20 unique English songs by 12 different people. The CMUDict was used
@@
 type for training AST models due to this level of detail. It also includes a mix of slow to fast
-melodies annd a balanced gender distribution \cite{NUSDataset}.
+melodies annd a balanced gender distribution \citep{NUSDataset}.
 
 \subsubsection{Free Music Archive}
-The Free Music Archive (FMA) dataset is a dataset of 106,574 tracks with 161 genres \cite{FMA}. This
+The Free Music Archive (FMA) dataset is a dataset of 106,574 tracks with 161 genres \citep{FMA}. This
 dataset is not a good candidate for direclty training AST models because it does not contain any lyrics
@@
 VocalSet is a 10 hour dataset of a capella singing from 20 professional singers
-demonstrating a variety of singing techniques \cite{VocalSet}. This dataset is a good candidate for
+demonstrating a variety of singing techniques \citep{VocalSet}. This dataset is a good candidate for
 training Voice Activity Detection (VAD) models because it contains onset and offset timestamps for
@@
 Many datasets were considered but left out due to the lack of availablility. Some of the most
-popular datasets such as Mauch's Dataset \cite{mirex2021} and Hansen's Dataset \cite{Hansen} are
-not publicly available anymore. Newer datasets such as DALI \cite{DALI} and DAMP! \cite{DAMP} are
+popular datasets such as Mauch's Dataset \citep{mirex2021} and Hansen's Dataset \citep{Hansen} are
+not publicly available anymore. Newer datasets such as DALI \citep{DALI} and DAMP! \citep{DAMP} are
 are hidden behind institutional logins and require manually requesting access.
 
-While both the Mauch's Dataset and Hansen's Datasets are quite small (Mauch has 20 songs, Hansen has 9 \cite{mirex2021}), the
+While both the Mauch's Dataset and Hansen's Datasets are quite small (Mauch has 20 songs, Hansen has 9 \citep{mirex2021}), the
 DALI and DAMP! datasets are much larger. The DALI dataset in particular used a version of the
 Teacher Student technique mentioned in \ref{sec:TeacherStudent} to label 105 songs with timestamps
-for the word and phoneme level \cite{DALI}. The DAMP! dataset is even larger with 300x30x2 song
+for the word and phoneme level \citep{DALI}. The DAMP! dataset is even larger with 300x30x2 song
 dataset. Both of these datasets would be excellent candidates for training AST models from their
@@
 and generate a sequence of phonemes using a Hidden Markov Model (HMM). This is the traditional
-approach to ASR and is also used in AST \cite{Hansen}.
+approach to ASR and is also used in AST \citep{Hansen}.
 
@@
 \subsubsection{wav2vec 2.0}
-One of the most recent breakthroughs in ASR is wav2vec 2.0 \cite{wav2vec}. This model uses a
+One of the most recent breakthroughs in ASR is wav2vec 2.0 \citep{wav2vec}. This model uses a
 self-supervised learning approach was used during the training of the model. Unlabeled audio was
 fed into the model to learn discrete speech units. These discrete speech units required the
-Gumbel-Softmax \cite{gumbelSoftmax} to allow for backpropagation. The model was then
-fine-tuned with a linear layer and CTC loss on labeled data to perform ASR \cite{wav2vec}.
+Gumbel-Softmax \citep{gumbelSoftmax} to allow for backpropagation. The model was then
+fine-tuned with a linear layer and CTC loss on labeled data to perform ASR \citep{wav2vec}.
 
-This approach was able to achieve state-of-the-art results on the LibriSpeech dataset \cite{Librispeech}
-and the TIMIT dataset \cite{TIMIT}. However, the thing that makes this approach the most promising
+This approach was able to achieve state-of-the-art results on the LibriSpeech dataset \citep{Librispeech}
+and the TIMIT dataset \citep{TIMIT}. However, the thing that makes this approach the most promising
 is the fact that after pre-training on a large amount of unlabeled data, the model can be fine-tuned
-on a 10 minute subset of labeled data to achieve 5.2 WER on the LibriSpeech clean dataset \cite{wav2vec}.
+on a 10 minute subset of labeled data to achieve 5.2 WER on the LibriSpeech clean dataset \citep{wav2vec}.
 This is very important for AST because there are so few labeled datasets available.
@@
 state-of-the-art results on the Jamendo dataset as well as on the DALI, Hansen, Mauch,
-and DAMP! datasets \cite{wav2vecTransfer}. The approach they performed did not change the first part
+and DAMP! datasets \citep{wav2vecTransfer}. The approach they performed did not change the first part
 of the model that generated the discrete speech units. Instead, they changed the last part of the
 model to have another branch that outputs the probability of the current word given previous words
-and context. This approach achieved a 33.13 WER on the Jamendo dataset \cite{wav2vecTransfer}.
+and context. This approach achieved a 33.13 WER on the Jamendo dataset \citep{wav2vecTransfer}.
 
@@
 audio of their model as similar to the LibriSpeech dataset as possible. To do this, they used
-Demucs v3 \cite{Demux} to separate the vocals from the accompaniment. They did not futher remove any
+Demucs v3 \citep{Demucs} to separate the vocals from the accompaniment. They did not futher remove any
 noise or singing specific features from the audio. For their labeled datasets, they used
@@
 
-\bibliographystyle{apalike}
-\bibliography{references}
+% \bibliographystyle{apalike}
+% \bibliography{references}
+\printbibliography
 \end{multicols*}

### 2023-12-02 report/paper.tex
more on whisperx and sort bib

@@
 
-\subsection{Whisper Word-Level Alignment}
+\subsection{Whisper and WhisperX Word-Level Alignment}
+
+\subsubsection{Whisper}
+Whisper a ASR model trained on a extremely large dataset of 680,000 hours of audio \citep{whisper}.
+Instead of using self-supervised learning like in wav2vec 2.0, Whisper uses a weakly supervised
+approach. This approach uses a large dataset of audio and text that is not high quality as many of
+the smaller datasets. Instead the dataset was generated by scraping the internet for audio and
+their corresponding transcriptions. Although Whisper did not improve the state-of-the-art on the
+LibriSpeech dataset when compared against wav2vec 2.0, did achieve improvements and therefore
+state-of-the-art results on the a variety of other speech datasets \citep{whisper}. The model
+performs well in zero-shot scenarios so it can achieve good results without fine-tuning on the
+target dataset \citep{whisper}. This is important for AST because there are so many genres and
+types of music and it would be difficult to fine-tune on all of them.
+
+\subsubsection{WhisperX}
+The original Whisper model was also able to provide some timestamps but they were not the focus of
+the paper. It was more of a proof of concept that the same model can be used for multiple tasks
+\citep{whisper}. WhisperX is a model that was trained specifically for word-level alignment and
+the speedup of the Whisper model \citep{whisperX}.
+
+On the TED-LIUM dataset, a dataset of TED talks \cite{tedlium3}, WhisperX was able to acheive a
+11.8x speedup over Whisper. This is comparable to wav2vec 2.0 which was a 10.3x speedup over
+Whisper. However, WhisperX was also able to achieve a WER of 9.7 compared to the 10.5 WER from
+Whisper and 19.8 WER from wav2vec 2.0 \citep{whisperX}. This is a significant improvement over
+of speed with a slight increase in performance. This is important for AST because it would allow
+for faster training and inference times.
+
+For the word alignment task, WhisperX uses an external phoneme alignment approach to generate
+word level timestamps using phoneme level timestamps \citep{whisperX}. The external phoneme
+alignment model was actually a wav2vec 2.0 model that used Dynamic Time Warping (DTW) to align
+the phonemes \citep{whisperX}. The first and last phonemes of each word were used to generate
+the word level timestamps \citep{whisperX}.
+
+The most important improvement of WhisperX over Whisper is VAD Cut \& Merge technique. This
+technique uses a Voice Activity Detection (VAD) model to cut the audio into segments at low voice
+activity sections. This decreases the hallucination of Whisper if a word was partly spoken in a
+audio segment which lead to the slight increase in WER \citep{whisperX}. The speedup of WhisperX
+was also due to this technique it allowed parallel processing of the audio segments where
+Whisper had to process sequenctially \citep{whisperX}.
 
@@
 
-% \bibliographystyle{apalike}
-% \bibliography{references}
 \printbibliography
+
 \end{multicols*}
+
 \end{document}

### 2023-12-02 report/paper.tex
updated to before training

@@
 achieve a 77.8\% Word Error Rate (WER) which still leaves a lot of room for improvement.
-{
-\color{red}
-Remember to record the performance of the state of the art models on this dataset
-}
-% \subsubsection{Children's Songs Dataset} All Korean :(
+
+\subsubsection{Children's Songs Dataset}
+This is a dataset of 50 English and 50 Korean children's songs with grapheme level timestamps and
+pitches. This dataset was sung by a single professional Korean singer in two separate keys for each
+song \citep{ChildrensSongs}. This dataset is a good candidate for training AST models because it
+includes the onset and offset information, pitch, and grapheme level timestamps that could be
+preprocessed to phoneme or word level timestamps. It is also a very clean dataset because of the
+professional performance and setting. The downside of this dataset is that graphemes are
+not as popular as phonemes, word or utterance labels in either ASR or AST.
+
 \subsubsection{MUSDB18}
@@
 
-While both the Mauch's Dataset and Hansen's Datasets are quite small (Mauch has 20 songs, Hansen has 9 \citep{mirex2021}), the
-DALI and DAMP! datasets are much larger. The DALI dataset in particular used a version of the
-Teacher Student technique mentioned in \ref{sec:TeacherStudent} to label 105 songs with timestamps
-for the word and phoneme level \citep{DALI}. The DAMP! dataset is even larger with 300x30x2 song
-dataset. Both of these datasets would be excellent candidates for training AST models from their
-size alone.
+While both the Mauch's Dataset and Hansen's Datasets are quite small (Mauch has 20 songs, Hansen has
+9 \citep{mirex2021}), the DALI and DAMP! datasets are much larger. The DALI dataset in particular
+used a version of the Teacher Student technique mentioned in \ref{sec:TeacherStudent} to label 105
+songs with timestamps for the word and phoneme level \citep{DALI}. The DAMP! dataset is even larger
+with 300x30x2 song dataset. Both of these datasets would be excellent candidates for training AST
+models from their size alone.
 
@@
 \color{red}
+Other models focus on the transcription of notes instead of lyrics. These models use a similar
+to the ones that are transcribing lyrics
 TODO: Add more detail about HMMs
@@
 word level timestamps using phoneme level timestamps \citep{whisperX}. The external phoneme
-alignment model was actually a wav2vec 2.0 model that used Dynamic Time Warping (DTW) to align
-the phonemes \citep{whisperX}. The first and last phonemes of each word were used to generate
-the word level timestamps \citep{whisperX}.
+alignment model was actually a wav2vec 2.0 model that used Dynamic Time Warping (DTW) \cite{DTW}
+to align the phonemes \citep{whisperX}. The first and last phonemes of each word were used to
+generate the word level timestamps \citep{whisperX}.
 
@@
 \section{Method}
-\begin{enumerate}
-    \item Preprocess datasets
-    \item Augment datasets
-    \item Create frankenstein dataset
-    \item Train model
-    \item Evaluate model
-    \item Label unlabeled datasets
-    \item train student model on newly labeled datasets
-    \item evaluate student model on original, manually labeled datasets
-    \item repeat
-\end{enumerate}
+In order to produce a model that can perform AST, we train a model to first take the audio and align
+it with the lyrics like in the last step of the WhisperX pipeline. This helps us create a model that
+can poperly annotate phonemes from the audio. Then, a secondary model can be trained to take the
+audio and the phoneme annotations to generate the lyrics. With both a lyric and phoneme generator,
+any song will be able to be transcribed into lyrics and time-aligned phonemes for an end-to-end AST.
+
+Each step in this process allows for the use of teacher-student techniques mentioned in section
+\ref{sec:TeacherStudent} to generate more labeled data from unlabeled data. This is important
+because there are so few labeled datasets available for AST.
+
+\subsection{Preprocessing Datasets}
+To preprocess datasets, the audio files were preprocessed in the same way WhisperX preprocessed
+data with the VAD Cut \& Merge technique \citep{whisperX} except with a slight modification before
+trying to detect voices. Since the audio files were songs, most of them will have both singing and
+background music. This means that the VAD model will have a harder time detecting when the singing
+stops. To help the VAD model, the audio files were first passed through Demucs v4 \citep{Demucs}
+to separate the vocals from the background music. Then, the VAD model was used to detect the
+voices. Finally, the audio files were cut into segments at the detected voice sections. To reduce
+memory requirements and because lyrics utterances are often shorter than 30 seconds, the segments
+will be cut into 8-15 second segments. Segments that are shorter than 8 seconds can either be padded
+or merged with other segments into segments of 8-15 seconds.
+
+The first subsystem of our AST model will be to take the audio and generate the phoneme annotations.
+This subsystem will be trained and tested on the TIMIT, Children's Song and NUS datasets. These are
+all datasets that contain phoneme level annotations. TIMIT, although it is a speech dataset, can be
+transformed into a pseudo-singing dataset as mentioned in \ref{sec:pseudoSinging} or used directly
+as a different variety of vocalizing phonemes.
+
+\subsection{Augmenting Datasets}
+There are many ways to augment datasets. The three ways mentioned in SpecAugment includes
+time warping, frequency masking and time masking \citep{SpecAugment}. These techniques can be
+applied to the audio spectrogram before it is fed into the model. Time warping is a technique that
+stretches or compresses the audio spectrogram in the time dimension so it must be handled carefully
+to update the labels accodingly. Another way to augment the datasets is to transform the audio
+pitch. This can be done by shifting the pitch up or down.
+
+% \subsection{Combining Datasets}
+
+\section{Training}
+
+% \begin{enumerate}
+%     \item Preprocess datasets
+%     \item Augment datasets
+%     \item Create frankenstein dataset
+%     \item Train model
+%     \item Evaluate model
+%     \item Label unlabeled datasets
+%     \item train student model on newly labeled datasets
+%     \item evaluate student model on original, manually labeled datasets
+%     \item repeat
+% \end{enumerate}
 
@@
 
-\subsection{WER}
+\subsection{VAD Performance}
 
-\subsection{PER}
+\subsection{Phoneme Alignment Performance}
+% PER
+
+\subsection{Lyric Generation Performance}
+% WER
 
@@
 remove pitch from singing
+\subsection{Phoneme to Grapheme}
+\citep{G2P} or \citep{P2G}
 
 \section{Conclusion}
-
+\citep{pyannote}
 \printbibliography

### 2023-12-02 data/test.py
add utils

@@ -1,0 +1,38 @@
+import unittest
+from .utils import *
+
+class UtilsTestCase(unittest.TestCase):
+
+    OUTPUT_DIR = Path("data/test_output")
+    # https://ccmixter.org/files/admiralbob77/65751
+    SAMPLE_MP3 = Path("data/admiralbob77_-_Creator_of_the_Stars_At_Night_3.mp3")
+
+    @classmethod
+    def setUpClass(cls):
+        cls.OUTPUT_DIR.mkdir(parents=True, exist_ok=True)
+
+    @classmethod
+    def tearDownClass(cls):
+        cls.OUTPUT_DIR.rmdir()
+
+    def testDemucs(self):
+        separate_vocals(
+            filepath=[self.SAMPLE_MP3],
+            output_dir=self.OUTPUT_DIR,
+            should_speedup=False,
+            shifts=1,
+            parallel=1,
+        )
+        filename = self.SAMPLE_MP3.stem
+        output_dir = Path(f"{self.OUTPUT_DIR}/{filename}")
+        vocals_path = Path(f"{output_dir}/vocals.wav")
+        no_vocals_path = Path(f"{output_dir}/no_vocals.wav")
+        self.assertTrue(vocals_path.exists())
+        self.assertTrue(no_vocals_path.exists())
+        vocals_path.unlink()
+        no_vocals_path.unlink()
+        output_dir.rmdir()
+
+
+if __name__ == "__main__":
+    unittest.main()

### 2023-12-02 data/utils.py

@@ -1,0 +1,26 @@
+from pathlib import Path
+import demucs.separate
+
+def separate_vocals(
+        filepath: list[Path],
+        output_dir: Path,
+        should_speedup: bool = False,
+        shifts: int = 1,
+        parallel: int = 1,
+    ):
+    """Separate vocals from a given audio file and save them in the output directory.
+
+    Args:
+        filepath (Path): Path to the audio file.
+        output_dir (Path): Path to the output directory.
+    """
+    output_dir.mkdir(parents=True, exist_ok=True)
+    arguments = [
+        "-n", f"{'htdemucs' if should_speedup else 'htdemucs_ft'}",
+        "-j", f"{parallel}",
+        "--shifts", f"{shifts}",
+        "--two-stems", "vocals",
+        "--filename", f"{output_dir.absolute()}/" + "{track}/{stem}.{ext}",
+        " ".join(f'{str(path)}' for path in filepath),
+    ]
+    demucs.separate.main(arguments)

### 2023-12-03 data/test.py
bad vad cuts

@@ -6,6 +6,7 @@ class UtilsTestCase(unittest.TestCase):
     OUTPUT_DIR = Path("data/test_output")
     # https://ccmixter.org/files/admiralbob77/65751
     SAMPLE_MP3 = Path("data/admiralbob77_-_Creator_of_the_Stars_At_Night_3.mp3")
+    vocals_path = None
 
     @classmethod
     def setUpClass(cls):
@@ -13,8 +14,12 @@ class UtilsTestCase(unittest.TestCase):
 
     @classmethod
     def tearDownClass(cls):
-        cls.OUTPUT_DIR.rmdir()
+        if False and cls.vocals_path and cls.vocals_path.exists():
+            cls.vocals_path.unlink()
+        if False and cls.OUTPUT_DIR.exists():
+            cls.OUTPUT_DIR.rmdir()
 
+    # @unittest.skip("Skip testDemucs")
     def testDemucs(self):
         separate_vocals(
             filepath=[self.SAMPLE_MP3],
@@ -29,9 +34,22 @@ class UtilsTestCase(unittest.TestCase):
         no_vocals_path = Path(f"{output_dir}/no_vocals.wav")
         self.assertTrue(vocals_path.exists())
         self.assertTrue(no_vocals_path.exists())
-        vocals_path.unlink()
         no_vocals_path.unlink()
+        return vocals_path
+
+    def testVadCut(self):
+        vocals_path = self.testDemucs()
+        filename = self.SAMPLE_MP3.stem
+        output_dir = Path(f"{self.OUTPUT_DIR}/cut/{filename}")
+        vad_cuts = vad_cut(
+            audio_filepath=vocals_path,
+            output_dir=self.OUTPUT_DIR,
+            max_duration_sec=10,
+            max_silence_sec=2,
+        )
+        self.assertTrue(output_dir.exists())
         output_dir.rmdir()
+        output_dir.parent.rmdir()
 
 
 if __name__ == "__main__":

### 2023-12-03 data/utils.py

@@ -1,5 +1,10 @@
 from pathlib import Path
+from pyannote.audio import Pipeline
+from .env import HF_TOKEN
+from dataclasses import dataclass
 import demucs.separate
+import os
+import torch
 
 def separate_vocals(
         filepath: list[Path],
@@ -24,3 +29,53 @@ def separate_vocals(
         " ".join(f'{str(path)}' for path in filepath),
     ]
     demucs.separate.main(arguments)
+
+@dataclass
+class VadCut:
+    filepath: Path
+    dataset: str
+    duration_ms: int
+
+@dataclass
+class TimedAnnotation:
+    start: int
+    end: int
+    label: str
+
+    def duration(self) -> int:
+        return self.end - self.start
+
+def vad_cut(audio_filepath: Path, output_dir: Path, max_duration_sec: float, max_silence_sec: float) -> list[VadCut]:
+    pipeline = Pipeline.from_pretrained(
+        "pyannote/speaker-diarization-3.1",
+        use_auth_token=HF_TOKEN,
+    )
+    pipeline.to(torch.device("cuda" if torch.cuda.is_available() else "cpu"))
+    print("========== Audio File", audio_filepath)
+    diarization = pipeline(audio_filepath)
+    valid_segements = []
+    for segment, label in diarization.itertracks(yield_label=False):
+        if len(valid_segements) == 0:
+            valid_segements.append(TimedAnnotation(segment.start, segment.end, label))
+        else:
+            last_segment = valid_segements[-1]
+            segment_duration_sec = segment.end - segment.start
+            if last_segment.duration() + segment_duration_sec < max_duration_sec \
+                    and segment.start - last_segment.end < max_silence_sec:
+                last_segment.end = segment.end
+            else:
+                valid_segements.append(TimedAnnotation(segment.start, segment.end, label))
+
+    valid_cuts = []
+    for segment in valid_segements:
+        segment_filepath = Path(f"{output_dir}/{audio_filepath.stem}_{segment.start}_{segment.end}.wav")
+        segment_filepath.parent.mkdir(parents=True, exist_ok=True)
+        os.system(f"ffmpeg -i {audio_filepath} -ss {segment.start} -to {segment.end} -c copy {segment_filepath}")
+        valid_cuts.append(VadCut(segment_filepath, segment.label, segment.duration()*1000))
+    print(f"Valid cuts: {len(valid_cuts)}")
+    return valid_cuts
+
+
+
+
+

### 2023-12-03 data/test.py
vad segments properly

@@ -6,7 +6,7 @@ class UtilsTestCase(unittest.TestCase):
     OUTPUT_DIR = Path("data/test_output")
     # https://ccmixter.org/files/admiralbob77/65751
     SAMPLE_MP3 = Path("data/admiralbob77_-_Creator_of_the_Stars_At_Night_3.mp3")
-    vocals_path = None
+
 
     @classmethod
     def setUpClass(cls):
@@ -14,9 +14,7 @@ class UtilsTestCase(unittest.TestCase):
 
     @classmethod
     def tearDownClass(cls):
-        if False and cls.vocals_path and cls.vocals_path.exists():
-            cls.vocals_path.unlink()
-        if False and cls.OUTPUT_DIR.exists():
+        if cls.OUTPUT_DIR.exists():
             cls.OUTPUT_DIR.rmdir()
 
     # @unittest.skip("Skip testDemucs")
@@ -40,15 +38,19 @@ class UtilsTestCase(unittest.TestCase):
     def testVadCut(self):
         vocals_path = self.testDemucs()
         filename = self.SAMPLE_MP3.stem
-        output_dir = Path(f"{self.OUTPUT_DIR}/cut/{filename}")
+        output_dir = Path(f"{self.OUTPUT_DIR}/{filename}/cut")
+        pipeline = get_voice_activity_segments()
         vad_cuts = vad_cut(
+            pipeline=pipeline,
             audio_filepath=vocals_path,
-            output_dir=self.OUTPUT_DIR,
-            max_duration_sec=10,
-            max_silence_sec=2,
+            output_dir=output_dir,
         )
         self.assertTrue(output_dir.exists())
+        for cut in vad_cuts:
+            self.assertTrue(cut.filepath.exists())
+            cut.filepath.unlink()
         output_dir.rmdir()
+        vocals_path.unlink()
         output_dir.parent.rmdir()
 
 

### 2023-12-03 data/utils.py

@@ -1,7 +1,9 @@
 from pathlib import Path
-from pyannote.audio import Pipeline
+from typing import Optional
+from pyannote.audio import Pipeline, Model, Inference
 from .env import HF_TOKEN
 from dataclasses import dataclass
+from .whisperX.vad import load_vad_model, VoiceActivitySegmentation, merge_chunks
 import demucs.separate
 import os
 import torch
@@ -45,29 +47,29 @@ class TimedAnnotation:
     def duration(self) -> int:
         return self.end - self.start
 
-def vad_cut(audio_filepath: Path, output_dir: Path, max_duration_sec: float, max_silence_sec: float) -> list[VadCut]:
-    pipeline = Pipeline.from_pretrained(
-        "pyannote/speaker-diarization-3.1",
-        use_auth_token=HF_TOKEN,
-    )
-    pipeline.to(torch.device("cuda" if torch.cuda.is_available() else "cpu"))
-    print("========== Audio File", audio_filepath)
-    diarization = pipeline(audio_filepath)
-    valid_segements = []
-    for segment, label in diarization.itertracks(yield_label=False):
-        if len(valid_segements) == 0:
-            valid_segements.append(TimedAnnotation(segment.start, segment.end, label))
-        else:
-            last_segment = valid_segements[-1]
-            segment_duration_sec = segment.end - segment.start
-            if last_segment.duration() + segment_duration_sec < max_duration_sec \
-                    and segment.start - last_segment.end < max_silence_sec:
-                last_segment.end = segment.end
-            else:
-                valid_segements.append(TimedAnnotation(segment.start, segment.end, label))
+def get_voice_activity_segments() -> VoiceActivitySegmentation:
+    device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
+    return load_vad_model(device, vad_offset=0.5)
+
+def vad_cut(
+        pipeline: VoiceActivitySegmentation,
+        audio_filepath: Path,
+        output_dir: Path,
+        max_duration_sec: float=10,
+        onset: float = 0.5,
+        offset: Optional[float] = None,
+    ) -> list[VadCut]:
+
+    segments = pipeline(audio_filepath)
+    output = merge_chunks(segments, max_duration_sec, onset=onset, offset=offset)
+
+    valid_segments = []
+    for segment in output:
+        print(f'{segment["end"] - segment["start"]:0.3f}', segment)
[…]

### 2023-12-03 data/whisperX/audio.py

@@ -1,0 +1,162 @@
+import os
+import subprocess
+from functools import lru_cache
+from typing import Optional, Union
+
+import numpy as np
+import torch
+import torch.nn.functional as F
+
+# from .utils import exact_div
+def exact_div(x, y):
+    assert x % y == 0
+    return x // y
+
+# hard-coded audio hyperparameters
+SAMPLE_RATE = 16000
+N_FFT = 400
+HOP_LENGTH = 160
+CHUNK_LENGTH = 30
+N_SAMPLES = CHUNK_LENGTH * SAMPLE_RATE  # 480000 samples in a 30-second chunk
+N_FRAMES = exact_div(N_SAMPLES, HOP_LENGTH)  # 3000 frames in a mel spectrogram input
+
+N_SAMPLES_PER_TOKEN = HOP_LENGTH * 2  # the initial convolutions has stride 2
+FRAMES_PER_SECOND = exact_div(SAMPLE_RATE, HOP_LENGTH)  # 10ms per audio frame
+TOKENS_PER_SECOND = exact_div(SAMPLE_RATE, N_SAMPLES_PER_TOKEN)  # 20ms per audio token
+
+
+def load_audio(file: str, sr: int = SAMPLE_RATE):
+    """
+    Open an audio file and read as mono waveform, resampling as necessary
+
+    Parameters
+    ----------
+    file: str
+        The audio file to open
+
+    sr: int
+        The sample rate to resample the audio if necessary
+
+    Returns
+    -------
+    A NumPy array containing the audio waveform, in float32 dtype.
+    """
+    try:
+        # Launches a subprocess to decode audio while down-mixing and resampling as necessary.
+        # Requires the ffmpeg CLI to be installed.
+        cmd = [
+            "ffmpeg",
+            "-nostdin",
+            "-threads",
+            "0",
+            "-i",
+            file,
+            "-f",
+            "s16le",
+            "-ac",
+            "1",
+            "-acodec",
+            "pcm_s16le",
+            "-ar",
+            str(sr),
+            "-",
+        ]
+        out = subprocess.run(cmd, capture_output=True, check=True).stdout
+    except subprocess.CalledProcessError as e:
+        raise RuntimeError(f"Failed to load audio: {e.stderr.decode()}") from e
+
+    return np.frombuffer(out, np.int16).flatten().astype(np.float32) / 32768.0
+
+
+def pad_or_trim(array, length: int = N_SAMPLES, *, axis: int = -1):
+    """
[…]

### 2023-12-03 data/test.py
updates vad to not generate new audio files

@@ -36,22 +36,15 @@ class UtilsTestCase(unittest.TestCase):
         return vocals_path
 
     def testVadCut(self):
-        vocals_path = self.testDemucs()
-        filename = self.SAMPLE_MP3.stem
-        output_dir = Path(f"{self.OUTPUT_DIR}/{filename}/cut")
         pipeline = get_voice_activity_segments()
         vad_cuts = vad_cut(
             pipeline=pipeline,
-            audio_filepath=vocals_path,
-            output_dir=output_dir,
+            audio_filepath=self.SAMPLE_MP3,
         )
-        self.assertTrue(output_dir.exists())
+        file = vad_cuts[0].filepath
         for cut in vad_cuts:
             self.assertTrue(cut.filepath.exists())
-            cut.filepath.unlink()
-        output_dir.rmdir()
-        vocals_path.unlink()
-        output_dir.parent.rmdir()
+            self.assertTrue(cut.filepath == file)
 
 
 if __name__ == "__main__":

### 2023-12-03 data/utils.py

@@ -1,11 +1,8 @@
 from pathlib import Path
 from typing import Optional
-from pyannote.audio import Pipeline, Model, Inference
-from .env import HF_TOKEN
 from dataclasses import dataclass
 from .whisperX.vad import load_vad_model, VoiceActivitySegmentation, merge_chunks
 import demucs.separate
-import os
 import torch
 
 def separate_vocals(
@@ -54,7 +51,7 @@ def get_voice_activity_segments() -> VoiceActivitySegmentation:
 def vad_cut(
         pipeline: VoiceActivitySegmentation,
         audio_filepath: Path,
-        output_dir: Path,
+        dataset="unknown",
         max_duration_sec: float=10,
         onset: float = 0.5,
         offset: Optional[float] = None,
@@ -62,19 +59,10 @@ def vad_cut(
 
     segments = pipeline(audio_filepath)
     output = merge_chunks(segments, max_duration_sec, onset=onset, offset=offset)
-
-    valid_segments = []
-    for segment in output:
-        print(f'{segment["end"] - segment["start"]:0.3f}', segment)
-        valid_segments.append(TimedAnnotation(segment["start"], segment["end"], segment["segments"]))
-
     valid_cuts = []
-    for segment in valid_segments:
-        segment_filepath = Path(f"{output_dir}/{audio_filepath.stem}_{segment.start}_{segment.end}.wav")
-        segment_filepath.parent.mkdir(parents=True, exist_ok=True)
-        os.system(f"ffmpeg -i {audio_filepath} -ss {segment.start} -to {segment.end} -c copy {segment_filepath}")
-        valid_cuts.append(VadCut(segment_filepath, segment.label, segment.duration()*1000))
-    print(f"Valid cuts: {len(valid_cuts)}")
+    for segment in output:
+        annotation = TimedAnnotation(segment["start"], segment["end"], segment["segments"])
+        valid_cuts.append(VadCut(audio_filepath, dataset, annotation.duration()*1000))
     return valid_cuts
 
 

### 2023-12-03 data/CMUdict/utils.py
add cmudict

@@ -1,0 +1,29 @@
+from dataclasses import dataclass
+from pathlib import Path
+
+FILE_DIR = Path(__file__).parent
+
+@dataclass
+class CMUDict:
+    def __init__(self):
+        self.map = {}
+
+        ## Load CMUdict
+        with open(FILE_DIR/"cmudict-0.7b.txt", "r") as f:
+            for line in f:
+                if line.startswith(";;;"):
+                    continue
+                word, phonemes = line.split("  ")
+                self.map[word] = phonemes.split()
+
+
+    def get_phonemes(self, word: str) -> list[str]:
+        """Get phonemes for a given word.
+
+        Args:
+            word (str): The word to get phonemes for.
+
+        Returns:
+            list[str]: The phonemes for the given word.
+        """
+        return self.map[word.upper()]

### 2023-12-03 data/csd/csd.py
add csd data creation

@@ -1,0 +1,178 @@
+#%% This directory should contain the english folder of the dataset.
+from dataclasses import dataclass
+import json
+import pandas as pd
+import subprocess
+import numpy as np
+from tqdm import tqdm
+import sys
+from pathlib import Path
+FILE_DIR = Path(__file__).parent
+DATA_DIR = FILE_DIR.parent
+if str(DATA_DIR) not in sys.path:
+    sys.path.append(str(DATA_DIR))
+else:
+    print("DATA_DIR already in sys.path", sys.path)
+from constants import *
+np.random.seed(42)
+# %%
+metadata = pd.read_json(FILE_DIR/"english/metadata.json", orient="index")
+metadata.head()
+
+#%%
+mapping = pd.read_csv(FILE_DIR/"phoneme-translation.csv")
+csd2cmu = dict(zip(mapping["csd"], mapping["cmu"]))
+csd2cmu
+
+# %%
+@dataclass
+class CSD_Song:
+    id: str
+    song_name: str
+
+    def get_csv(self) -> pd.DataFrame:
+        return pd.read_csv(FILE_DIR / f"english/csv/{self.id}.csv")
+
+    def get_lyrics(self) -> list[str]:
+        lyric_lines = []
+        with open(FILE_DIR / f"english/lyric/{self.id}.txt", "r") as f:
+            lyric_lines = f.readlines()
+        return [lyric.strip() for lyric in lyric_lines]
+
+    def get_syllables(self) -> list[str]:
+        syllables_lines = []
+        with open(FILE_DIR / f"english/txt/{self.id}.txt", "r") as f:
+            for syllables_line in f.readlines():
+                syllables_line = syllables_line.strip()
+                if syllables_line == "":
+                    continue
+                syllables_lines.append(syllables_line.split())
+        return syllables_lines
+
+    def get_csd_phonemes(self) -> list[str]:
+        phonemes_lines = []
+        for syllable_line in self.get_syllables():
+            current_phoneme_line = []
+            for syllable in syllable_line:
+                for csd_phoneme in syllable.split("_"):
+                    current_phoneme_line.append(csd_phoneme)
+            phonemes_lines.append(current_phoneme_line)
+        return phonemes_lines
+
+    def get_cmu_phonemes(self) -> list[str]:
+        cmu_phoneme_lines = []
+        for csd_phoneme in self.get_csd_phonemes():
+            cmu_phonemes = []
+            print(csd_phoneme)
[…]

### 2023-12-03 data/test.py
update test

@@ -17,7 +17,6 @@ class UtilsTestCase(unittest.TestCase):
         if cls.OUTPUT_DIR.exists():
             cls.OUTPUT_DIR.rmdir()
 
-    # @unittest.skip("Skip testDemucs")
     def testDemucs(self):
         separate_vocals(
             filepath=[self.SAMPLE_MP3],
@@ -33,7 +32,8 @@ class UtilsTestCase(unittest.TestCase):
         self.assertTrue(vocals_path.exists())
         self.assertTrue(no_vocals_path.exists())
         no_vocals_path.unlink()
-        return vocals_path
+        vocals_path.unlink()
+        output_dir.rmdir()
 
     def testVadCut(self):
         pipeline = get_voice_activity_segments()

### 2023-12-03 data/utils.py

@@ -65,7 +65,3 @@ def vad_cut(
         valid_cuts.append(VadCut(audio_filepath, dataset, annotation.duration()*1000))
     return valid_cuts
 
-
-
-
-

### 2023-12-03 data/csd/csd.py
update csd

@@ -170,6 +170,7 @@ test_valid = dataset["test"].train_test_split(test_size=0.5, seed=42)
 dataset["test"] = test_valid["test"]
 dataset["validation"] = test_valid["train"]
 dataset = dataset.cast_column("filename", Audio(sampling_rate=16000))
+dataset = dataset.rename_column("filename", "audio")
 dataset.save_to_disk(FILE_DIR/"csd_dataset")
 # %% Cleanup segments directory
 for file in (FILE_DIR/"segments").glob("*.wav"):

### 2023-12-03 data/timit/timit.py
add timit

@@ -1,0 +1,10 @@
+#%%
+import datasets
+from pathlib import Path
+
+#%%
+timit = datasets.load_dataset("timit_asr", data_dir=Path(__file__).parent.parent)
+timit
+# %%
+timit.save_to_disk(Path(__file__).parent / "timit_dataset")
+# %%

### 2023-12-03 data/test.py
update data utils

@@ -36,7 +36,7 @@ class UtilsTestCase(unittest.TestCase):
         output_dir.rmdir()
 
     def testVadCut(self):
-        pipeline = get_voice_activity_segments()
+        pipeline = get_voice_activity_pipeline()
         vad_cuts = vad_cut(
             pipeline=pipeline,
             audio_filepath=self.SAMPLE_MP3,

### 2023-12-03 data/utils.py

@@ -44,7 +44,7 @@ class TimedAnnotation:
     def duration(self) -> int:
         return self.end - self.start
 
-def get_voice_activity_segments() -> VoiceActivitySegmentation:
+def get_voice_activity_pipeline() -> VoiceActivitySegmentation:
     device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
     return load_vad_model(device, vad_offset=0.5)
 

### 2023-12-03 data/nus/nus.py
add nus

@@ -1,0 +1,167 @@
+#%%
+from dataclasses import dataclass
+from datasets import Dataset, Audio
+from pathlib import Path
+from tqdm import tqdm
+import subprocess
+import json
+FILE_DIR = Path(__file__).parent
+
+DATA_DIR = FILE_DIR/"data"
+import sys
+if str(FILE_DIR.parent) not in sys.path:
+    sys.path.append(str(FILE_DIR.parent))
+else:
+    print("DATA_DIR already in sys.path", sys.path)
+from utils import get_voice_activity_pipeline, vad_cut, VadCut
+
+#%%
+pipeline = get_voice_activity_pipeline()
+#%%
+@dataclass
+class NUS_Song:
+    singer: str
+    song: str
+    is_read: bool
+    is_sing: bool
+
+    @property
+    def read_sing_dir(self) -> str:
+        if self.is_read:
+            return "read"
+        elif self.is_sing:
+            return "sing"
+        else:
+            raise Exception("Unknown read_sing_dir name")
+
+    @property
+    def audio_file(self) -> Path:
+        return DATA_DIR/self.singer/self.read_sing_dir/f"{self.song}.wav"
+
+    @property
+    def phonemes(self) -> list[str]:
+        phonemes = []
+        with open(DATA_DIR/self.singer/self.read_sing_dir/f"{self.song}.txt", "r") as f:
+            for line in f.readlines():
+                start, end, phoneme = line.strip().split()
+                phonemes.append({
+                    "start": float(start),
+                    "end": float(end),
+                    "phoneme": phoneme.upper(),
+                })
+        return phonemes
+
+    def get_segements(self) -> list:
+        segments = []
+        vad_cuts = vad_cut(
+            pipeline=pipeline,
+            audio_filepath=self.audio_file,
+        )
+        for cut in vad_cuts:
+            segments.append(NUS_Segment(
+                song=self,
+                start=cut.annotation.start,
+                end=cut.annotation.end,
+            ))
+        return segments
+
+@dataclass
+class NUS_Segment:
+    song: NUS_Song
+    start: float
+    end: float
+
+    @property
+    def duration(self) -> float:
+        return self.end - self.start
+
+    @property
+    def segment_filename(self) -> Path:
+        path = FILE_DIR / f"segments/{self.song.singer}_{'read' if self.song.is_read else 'sing'}_{self.start}_{self.end}.wav"
+        if not path.parent.exists():
[…]

### 2023-12-03 data/utils.py
update utils

@@ -1,7 +1,10 @@
 from pathlib import Path
 from typing import Optional
 from dataclasses import dataclass
-from .whisperX.vad import load_vad_model, VoiceActivitySegmentation, merge_chunks
+import sys
+if str(Path(__file__).parent) not in sys.path:
+    sys.path.append(str(Path(__file__).parent))
+from whisperX.vad import load_vad_model, VoiceActivitySegmentation, merge_chunks
 import demucs.separate
 import torch
 
@@ -30,12 +33,6 @@ def separate_vocals(
     demucs.separate.main(arguments)
 
 @dataclass
-class VadCut:
-    filepath: Path
-    dataset: str
-    duration_ms: int
-
-@dataclass
 class TimedAnnotation:
     start: int
     end: int
@@ -43,6 +40,12 @@ class TimedAnnotation:
 
     def duration(self) -> int:
         return self.end - self.start
+
+@dataclass
+class VadCut:
+    filepath: Path
+    dataset: str
+    annotation: TimedAnnotation
 
 def get_voice_activity_pipeline() -> VoiceActivitySegmentation:
     device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
@@ -62,6 +65,6 @@ def vad_cut(
     valid_cuts = []
     for segment in output:
         annotation = TimedAnnotation(segment["start"], segment["end"], segment["segments"])
-        valid_cuts.append(VadCut(audio_filepath, dataset, annotation.duration()*1000))
+        valid_cuts.append(VadCut(audio_filepath, dataset, annotation))
     return valid_cuts
 

### 2023-12-04 data/CMUdict/utils.py
add jamendo lyrics

@@ -3,11 +3,49 @@ from pathlib import Path
 
 FILE_DIR = Path(__file__).parent
 
+additional_words = {
+    "WASTIN'": "W EY1 S T IH0 NG".split(),
+    "LA": "L AA1".split(),
+    "LALALALALA": "L AA1 L AA1 L AA1 L AA1".split(),
+    "LALALALA": "L AA1 L AA1 L AA1 L AA1".split(),
+    "GETTING'":  "G IH1 T AH0 N".split(),
+    "PEYOTE": "P EY0 OW1 T IY0".split(),
+    "KNIFES": "N AY1 F".split(),
+    "BREATHIN'": "B R EH1 TH IH0 NG".split(),
+    "HUHHH": "HH AH1 N".split(),
+    "HUHHHH": "HH AH1 N".split(),
+    "WHUTSUP": "W AH1 T S AH2 P".split(),
+    "NIGGAS": "N IH1 G AH0 S".split(),
+    "NIGGA": "N IH1 G AH0".split(),
+    "HOMIE": "HH OW1 M IY0".split(),
+    "WORDLESSLY": "W ER1 D L AH0 S l IY0".split(),
+    "REPPIN": "R EH1 P IH0 NG".split(),
+    "GOTCHU": "G AA1 CH UW1".split(),
+    "FAM": "F AH0 M".split(),
+    "D'YOU": "D Y UW1".split(),
+    "ACCURSED": "AH0 K ER1 S T".split(),
+    "UNPERSUADED": "AH2 N P ER0 S W EY1 D IH0 D".split(),
+}
+
+misspellings = {
+    "AINT": "AIN'T",
+    "COMPLETLY": "COMPLETELY",
+    "THATS": "THAT'S",
+    "SEPERATED": "SEPARATED",
+    "STOPPIN": "STOPPING",
+    "POPPIN": "POPPING",
+    "SLIPPIN": "SLIPPING",
+    "WASNT": "WASN'T",
+    "BELEIVING": "BELIEVING",
+    "DOIN": "DOING",
+    "PARLIMENT": "PARLIAMENT"
+}
+
 @dataclass
 class CMUDict:
     def __init__(self):
         self.map = {}
-
+        self.unknown = []
         ## Load CMUdict
         with open(FILE_DIR/"cmudict-0.7b.txt", "r") as f:
             for line in f:
@@ -15,6 +53,10 @@ class CMUDict:
                     continue
                 word, phonemes = line.split("  ")
[…]

### 2023-12-04 data/jamendolyrics/jamendolyrics.py

@@ -1,0 +1,154 @@
+#%%
+from dataclasses import dataclass
+from tqdm import tqdm
+from pathlib import Path
+from datasets import Dataset, Audio
+import subprocess
+import pandas as pd
+import json
+import sys
+FILE_DIR = Path(__file__).parent
+DATA_DIR = FILE_DIR.parent
+if str(DATA_DIR) not in sys.path:
+    sys.path.append(str(DATA_DIR))
+else:
+    print("DATA_DIR already in sys.path", sys.path)
+from CMUdict.utils import CMUDict
+#%%
+cmudict = CMUDict()
+#%%
+@dataclass
+class JamendoLyrics_Song:
+    artist: str
+    title: str
+    genre: str
+    language: str
+    filename: str
+
+    @property
+    def audio_file(self) -> Path:
+        return FILE_DIR/f"data/mp3/{self.filename}.mp3"
+
+    @property
+    def lyrics_file(self) -> Path:
+        return FILE_DIR/f"data/annotations/lines/{self.filename}.csv"
+
+    def get_segments(self, max_duration: float = 10.0):
+        segments = []
+        df = pd.read_csv(self.lyrics_file)
+        current_segments = []
+        current_lyrics = []
+        for _, row in df.iterrows():
+            previous_duration = 0.0
+            if len(current_segments) > 0:
+                previous_duration = current_segments[-1]['end_time'] - current_segments[0]['start_time']
+            current_duration = row["end_time"] - row["start_time"]
+            if previous_duration + current_duration >= max_duration:
+                segments.append(JamendoLyics_Segment(
+                    song=self,
+                    start=current_segments[0]['start_time'],
+                    end=current_segments[-1]['end_time'],
+                    lyrics=" ".join(current_lyrics),
+                ))
+                current_segments = []
+                current_lyrics = []
+            current_segments.append(row)
+            current_lyrics.append(row["lyrics_line"].strip())
+        if len(current_segments) > 0:
+            segments.append(JamendoLyics_Segment(
+                song=self,
+                start=current_segments[0]['start_time'],
+                end=current_segments[-1]['end_time'],
+                lyrics=" ".join(current_lyrics),
+            ))
+        return segments
+
+@dataclass
+class JamendoLyics_Segment:
+    song: JamendoLyrics_Song
+    start: float
+    end: float
+    lyrics: str
+
+    @property
[…]

### 2023-12-04 data/nus/nus.py
update the datasets

@@ -115,7 +115,8 @@ class NUS_Segment:
             "end": self.end,
             "duration": self.duration,
             "filename": str(self.segment_filename),
-            "phonemes": self.phonemes,
+            "timestamped_phonemes": self.phonemes,
+            "phonemes": [phoneme["phoneme"] for phoneme in self.phonemes],
         }
 #%%
 songs = []

### 2023-12-04 data/timit/timit.py

@@ -4,7 +4,11 @@ from pathlib import Path
 
 #%%
 timit = datasets.load_dataset("timit_asr", data_dir=Path(__file__).parent.parent)
-timit
 # %%
+train_test_split = timit["train"].train_test_split(test_size=0.2, seed=42)
+timit['train'] = train_test_split['train']
+timit['validation'] = train_test_split['test']
+timit
+#%%
 timit.save_to_disk(Path(__file__).parent / "timit_dataset")
 # %%

### 2023-12-04 code/alignment/train.py
can train

@@ -1,0 +1,5 @@
+import torch
+import torchaudio
+
+device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
+

### 2023-12-04 data/CMUdict/utils.py

@@ -3,6 +3,8 @@ from pathlib import Path
 
 FILE_DIR = Path(__file__).parent
 
+UNK = "[UNK]"
+PAD = "[PAD]"
 additional_words = {
     "WASTIN'": "W EY1 S T IH0 NG".split(),
     "LA": "L AA1".split(),
@@ -41,11 +43,31 @@ misspellings = {
     "PARLIMENT": "PARLIAMENT"
 }
 
+# TimitBet 61 phoneme mapping to 39 phonemes
+# by Lee, K.-F., & Hon, H.-W. (1989). Speaker-independent phone recognition using hidden Markov models. IEEE Transactions on Acoustics, Speech, and Signal Processing, 37(11), 1641–1648. doi:10.1109/29.46546
+phon61_map39 = {
+    'iy':'iy',  'ih':'ih',   'eh':'eh',  'ae':'ae',    'ix':'ih',  'ax':'ah',   'ah':'ah',  'uw':'uw',
+    'ux':'uw',  'uh':'uh',   'ao':'aa',  'aa':'aa',    'ey':'ey',  'ay':'ay',   'oy':'oy',  'aw':'aw',
+    'ow':'ow',  'l':'l',     'el':'l',  'r':'r',      'y':'y',    'w':'w',     'er':'er',  'axr':'er',
+    'm':'m',    'em':'m',     'n':'n',    'nx':'n',     'en':'n',  'ng':'ng',   'eng':'ng', 'ch':'ch',
+    'jh':'jh',  'dh':'dh',   'b':'b',    'd':'d',      'dx':'dx',  'g':'g',     'p':'p',    't':'t',
+    'k':'k',    'z':'z',     'zh':'sh',  'v':'v',      'f':'f',    'th':'th',   's':'s',    'sh':'sh',
+    'hh':'hh',  'hv':'hh',   'pcl':'sil', 'tcl':'sil', 'kcl':'sil', 'qcl':'sil','bcl':'sil','dcl':'sil',
+    'gcl':'sil','h#':'sil',  '#h':'sil',  'pau':'sil', 'epi': 'sil','nx':'n',   'ax-h':'ah','q':'sil',
+    'sil': 'sil', UNK: UNK,  PAD: PAD, "sp": "sil"
+}
+PHON61_MAP39 = {}
+for key, value in phon61_map39.items():
+    PHON61_MAP39[key.upper()] = value.upper()
+
+VOCAB = { phoneme: i  for i, phoneme in enumerate(set(PHON61_MAP39.values())) }
[…]

### 2023-12-04 data/jamendolyrics/jamendolyrics.py

@@ -117,38 +117,35 @@ class JamendoLyics_Segment:
 
 
 #%%
-df = pd.read_csv(FILE_DIR/"data/JamendoLyrics.csv")
-df = df[df["Language"] == "English"]
-df.head()
-
-# %%
-songs = []
-for _, row in tqdm(df.iterrows(), total=len(df)):
-    songs.append(JamendoLyrics_Song(
-        artist=row["Artist"],
-        title=row["Title"],
-        genre=row["Genre"],
-        language=row["Language"],
-        filename=row["Filepath"][:-4],
-    ))
-# %%
-segments = []
-for song in tqdm(songs):
-    for segment in song.get_segments():
-        segments.append(segment)
-segments[0]
-# %%
-with open(FILE_DIR/"jamendolyrics.jsonl", "w") as f:
-    for segment in tqdm(segments):
-        f.write(json.dumps(segment.json_dict) + "\n")
-print(cmudict.unknown)
+def create_jamendo_jsonl():
+    df = pd.read_csv(FILE_DIR/"data/JamendoLyrics.csv")
+    df = df[df["Language"] == "English"]
+    songs = []
+    for _, row in tqdm(df.iterrows(), total=len(df)):
+        songs.append(JamendoLyrics_Song(
+            artist=row["Artist"],
+            title=row["Title"],
+            genre=row["Genre"],
+            language=row["Language"],
+            filename=row["Filepath"][:-4],
+        ))
+    segments = []
+    for song in tqdm(songs):
+        for segment in song.get_segments():
+            segments.append(segment)
+    segments[0]
+    with open(FILE_DIR/"jamendolyrics.jsonl", "w") as f:
+        for segment in tqdm(segments):
+            f.write(json.dumps(segment.json_dict) + "\n")
 #%%
-dataset = Dataset.from_json(str(FILE_DIR/"jamendolyrics.jsonl"))
-dataset = dataset.cast_column("filename", Audio(sampling_rate=16_000)).rename_column("filename", "audio")
-dataset.save_to_disk(FILE_DIR/"jamendolyrics_dataset")
 
-# %% Cleanup segments directory
-for file in (FILE_DIR/"segments").glob("*.mp3"):
-    file.unlink()
-(FILE_DIR/"segments").rmdir()
-#%%
+def get_jamendo():
+    if not (FILE_DIR/"jamendolyrics.jsonl").exists():
+        create_jamendo_jsonl()
[…]

### 2023-12-04 data/nus/nus.py

@@ -14,9 +14,11 @@ if str(FILE_DIR.parent) not in sys.path:
 else:
     print("DATA_DIR already in sys.path", sys.path)
 from utils import get_voice_activity_pipeline, vad_cut, VadCut
+from CMUdict.utils import CMUDict
 
 #%%
 pipeline = get_voice_activity_pipeline()
+cmudict = CMUDict()
 #%%
 @dataclass
 class NUS_Song:
@@ -44,10 +46,11 @@ class NUS_Song:
         with open(DATA_DIR/self.singer/self.read_sing_dir/f"{self.song}.txt", "r") as f:
             for line in f.readlines():
                 start, end, phoneme = line.strip().split()
+                # print(self, start, end, phoneme)
                 phonemes.append({
                     "start": float(start),
                     "end": float(end),
-                    "phoneme": phoneme.upper(),
+                    "phoneme": cmudict.map_phoneme(phoneme),
                 })
         return phonemes
 
@@ -77,7 +80,7 @@ class NUS_Segment:
 
     @property
     def segment_filename(self) -> Path:
-        path = FILE_DIR / f"segments/{self.song.singer}_{'read' if self.song.is_read else 'sing'}_{self.start}_{self.end}.wav"
+        path = FILE_DIR / f"segments/{self.song.singer}_{self.song.song}_{'read' if self.song.is_read else 'sing'}_{self.start}_{self.end}.wav"
         if not path.parent.exists():
             path.parent.mkdir(parents=True, exist_ok=True)
         if not path.exists():
@@ -119,50 +122,59 @@ class NUS_Segment:
             "phonemes": [phoneme["phoneme"] for phoneme in self.phonemes],
         }
 #%%
-songs = []
-for singer in DATA_DIR.iterdir():
-    if singer.is_dir():
-        for read_sing_dir in singer.iterdir():
-            if read_sing_dir.is_dir():
-                items = set()
-                for file in read_sing_dir.iterdir():
-                    if file.is_file():
-                        items.add(file.stem)
-                for item in items:
-                    if read_sing_dir.name == "read":
-                        is_read = True
-                        is_sing = False
-                    elif read_sing_dir.name == "sing":
-                        is_read = False
-                        is_sing = True
-                    else:
-                        raise Exception("Unknown read_sing_dir name")
-                    song = NUS_Song(
-                        singer=singer.name,
-                        song=item,
-                        is_read=is_read,
[…]

### 2023-12-04 data/timit/timit.py

@@ -1,14 +1,36 @@
 #%%
 import datasets
 from pathlib import Path
+import sys
+if str(Path(__file__).parent.parent) not in sys.path:
+    sys.path.append(str(Path(__file__).parent.parent))
+from CMUdict.utils import CMUDict
 
-#%%
-timit = datasets.load_dataset("timit_asr", data_dir=Path(__file__).parent.parent)
+DEBUG = False
+def get_timit():
+    cmuDict = CMUDict()
+    timit = datasets.load_dataset("timit_asr", data_dir=Path(__file__).parent.parent)
+    train_test_split = timit["train"].train_test_split(test_size=0.2, seed=42)
+    timit['train'] = train_test_split['train']
+    timit['validation'] = train_test_split['test']
+    def get_phonemes(item):
+        phonemes = item['phonetic_detail']['utterance']
+        item['phonemes'] = [cmuDict.map_phoneme(phoneme) for phoneme in phonemes]
+        return item
+    timit = timit.map(get_phonemes).remove_columns([
+        'phonetic_detail',
+        'file',
+        'word_detail',
+        'dialect_region',
+        'id',
+        'sentence_type',
+        'speaker_id',
+    ])
+    if DEBUG:
+        print(len(cmuDict.seen), cmuDict.seen)
+    return timit
 # %%
-train_test_split = timit["train"].train_test_split(test_size=0.2, seed=42)
-timit['train'] = train_test_split['train']
-timit['validation'] = train_test_split['test']
-timit
-#%%
-timit.save_to_disk(Path(__file__).parent / "timit_dataset")
-# %%
+if __name__ == "__main__":
+    DEBUG = True
+    timit = get_timit()
+    timit

### 2023-12-05 report/paper.tex
add specaugment figures

@@
 
-\usepackage[%
+\usepackage[
 backend=biber,
@@
 \author{Victor Shan}
+
 \begin{document}
@@
 augment the audio spectrogram to improve the performance of ASR models. A frequent intermediate
-representation are Mel-Frequency Cepstral Coefficients (MFCC). These are image representations of
+representation are Mel-Frequency Cepstral Coefficients (MFCC).
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.4\textwidth]{assets/waveform-to-spectrogram-sing.png}
+    \includegraphics[width=0.4\textwidth]{assets/melspectrogram-mfcc.png}
+    \caption{Waveform (top left) can be transformed into a spectrogram (bottom left).
+    The spectrogram can be transformed to use the Mel scale in Mel Spectrogram (top right,
+    only first section shows) which can futher be transformed into a more discrete MFCC.
+    Figures inspired by torchaudio \citep{torchAudioFigures}.}
+    \label{fig:MFCC}
+\end{figure*}
+
+These are image representations of
 energy at certain frequencies on a scale that more closely matches human hearing \citep{MFCC}.
@@
 
-On the TED-LIUM dataset, a dataset of TED talks \cite{tedlium3}, WhisperX was able to acheive a
+On the TED-LIUM dataset, a dataset of TED talks \citep{tedlium3}, WhisperX was able to acheive a
 11.8x speedup over Whisper. This is comparable to wav2vec 2.0 which was a 10.3x speedup over
@@
 word level timestamps using phoneme level timestamps \citep{whisperX}. The external phoneme
-alignment model was actually a wav2vec 2.0 model that used Dynamic Time Warping (DTW) \cite{DTW}
+alignment model was actually a wav2vec 2.0 model that used Dynamic Time Warping (DTW) \citep{DTW}
 to align the phonemes \citep{whisperX}. The first and last phonemes of each word were used to

### 2023-12-05 report/paper.tex
minor organizational paper changes

@@
 
-\title{Automatic Singing Transcription}
+\title{Automatic Singing Transcription: Phonemes and Alignment}
 \author{Victor Shan}
@@
 way to do this is to augment the existing datasets. SpecAugment is a series of techniques that
-augment the audio spectrogram to improve the performance of ASR models. A frequent intermediate
-representation are Mel-Frequency Cepstral Coefficients (MFCC).
+augment the audio spectrogram to improve the performance of ASR models. A common variation of this
+representation is the Mel-Frequency Cepstral Coefficient (MFCC) spectrogram.
 
@@
 
-% ----------------- INCLUDE MFCC IMAGE -----------------
+{
+    \centering
+    \includegraphics[width=0.45\textwidth]{assets/specaugment.png}
+}
 
@@
 level datasets into phoneme level datasets. There are many ASR datasets that contain single utterances
-such as LibriSpeech and MUSDB18 \citep{Librispeech,musdb18}. These datasets can be effectively
+such as LibriSpeech and JamendoLyrics. These datasets can be effectively
 transformed into phoneme level datasets by using a phoneme dictionary such as the CMU Pronouncing
@@
     \includegraphics[width=0.5\textwidth]{assets/CTC.png}
-    \caption{CTC Function collapsing repetitions \citep{CTC}}
+    \caption{Example of CTC Function collapsing repetitions. Image from \citep{CTC}}
     \label{fig:CTC}
@@
 pitches. This dataset was sung by a single professional Korean singer in two separate keys for each
-song \citep{ChildrensSongs}. This dataset is a good candidate for training AST models because it
+song \citep{CSD}. This dataset is a good candidate for training AST models because it
 includes the onset and offset information, pitch, and grapheme level timestamps that could be
@@
 memory requirements and because lyrics utterances are often shorter than 30 seconds, the segments
-will be cut into 8-15 second segments. Segments that are shorter than 8 seconds can either be padded
-or merged with other segments into segments of 8-15 seconds.
+will be cut into 5-15 second segments. Smaller segments can either be padded
+or merged with other segments into larger ones.
 
@@
 applied to the audio spectrogram before it is fed into the model. Time warping is a technique that
-stretches or compresses the audio spectrogram in the time dimension so it must be handled carefully
-to update the labels accodingly. Another way to augment the datasets is to transform the audio
-pitch. This can be done by shifting the pitch up or down.
+stretches or compresses the audio spectrogram in the time dimension due to the complexity of
+realigning the audio. We use time and frequency maksing for training our models.
 
@@
 
-\subsection{VAD Performance}
+% \subsection{VAD Performance}
 
@@
 
-\subsection{Lyric Generation Performance}
-% WER
+% \subsection{Lyric Generation Performance}
+% % WER
 
@@
 \section{Future Work}
+\subsection{Train on More Datasets}
+\subsection{Finetuning Wav2Vec 2.0 Feature Extractor}
 \subsection{Adversarial Training}

### 2023-12-05 data/csd/csd.py
train csd

@@ -3,7 +3,6 @@ from dataclasses import dataclass
 import json
 import pandas as pd
 import subprocess
-import numpy as np
 from tqdm import tqdm
 import sys
 from pathlib import Path
@@ -13,8 +12,6 @@ if str(DATA_DIR) not in sys.path:
     sys.path.append(str(DATA_DIR))
 else:
     print("DATA_DIR already in sys.path", sys.path)
-from constants import *
-np.random.seed(42)
 # %%
 metadata = pd.read_json(FILE_DIR/"english/metadata.json", orient="index")
 metadata.head()
@@ -131,49 +128,48 @@ def get_segments(song: CSD_Song) -> list[CSD_Segment]:
     return segments
 
 #%%
-song = CSD_Song("en001a", "Alphabet")
-segments = get_segments(song)
-segments
+# song = CSD_Song("en001a", "Alphabet")
+# segments = get_segments(song)
+# segments
 
 # %%
-split = {
-    "train": 0.8,
-    "validation": 0.1,
-    "test": 0.1
-}
-id = 0
-with open(FILE_DIR/"csd.jsonl", "w") as f:
-    for song_id, row in tqdm(metadata.iterrows(), total=len(metadata)):
-        song = CSD_Song(song_id, row["songname"])
-        segments = get_segments(song)
-        for segment in segments:
-            segment_dict = {
-                "id": id,
-                "song_id": segment.song_id,
-                "song_name": segment.song_name,
-                "start": segment.start,
-                "end": segment.end,
-                "filename": str(segment.segment_filename),
-                "phonemes": " ".join(segment.cmu_phonemes),
-                "lyrics": segment.lyrics,
-            }
-            id += 1
-            f.write(f'{json.dumps(segment_dict)}\n')
-    f.flush()
-f.close()
+def create_csd_jsonl():
+    id = 0
+    with open(FILE_DIR/"csd.jsonl", "w") as f:
+        for song_id, row in tqdm(metadata.iterrows(), total=len(metadata)):
+            song = CSD_Song(song_id, row["songname"])
+            segments = get_segments(song)
+            for segment in segments:
+                segment_dict = {
+                    "id": id,
+                    "song_id": segment.song_id,
+                    "song_name": segment.song_name,
+                    "start": segment.start,
+                    "end": segment.end,
[…]

### 2023-12-05 report/paper.tex
add future works

@@
 transformed into phoneme level datasets by using a phoneme dictionary such as the CMU Pronouncing
-Dictionary \citep{CMUDict}. This dictionary contains a mapping from words to phonemes. Using this
+Dictionary \citep{CMUDict}. This dictionary contains a mapping from words to
+ARPAbet style phonemes \citep{CMUDict}. Using this
 dictionary, the utterances can be transformed into phoneme sequences. These sequences can then be
@@
 
-\subsection{Existing Datasets}
+\subsection{Existing Datasets} \label{sec:datasets}
 
@@
 
+In order to transform these datsets into a valid input for the model, datasets that only contain
+utterances will be transformed into phoneme level datasets using the CMU Pronouncing Dictionary
+(CMUDict) and then tokenized. Luckily, the TIMIT dataset already uses an extended version of the
+ARPAbet phonemes so they can be remapped into a smaller subset of phonemes. The CMUDict also
+includes stresses in the form of numbers at the end of some phonemes that were removed. Finally,
+the unknown and padding tokens were added to the phoneme vocabulary. Any words that appeared in
+the datasets but were not in CMUDict were manually added to the dictionary. HuggingFace and PyTorch
+constructs to help with this process and allow for batching.
+
 \subsection{Augmenting Datasets}
@@
 \section{Training}
+
+\subsection{Base Model}
+In order to get a baseline for our improvements, we train a base model on the TIMIT dataset. This
+This dataset is a one of the most popular and defacto standard for ASR models that do phoneme
+recognition. The starting model that will be finetuned on TIMIT will be the `facebook/wav2vec2-base-960h'
+model from the HuggingFace \citep{huggingFaceWav2vec2}.
+
+\subsection{Singer Model}
 
@@
 % \subsection{VAD Performance}
+\subsection{Phoneme Recognition Performance}
 
@@
 \subsection{Train on More Datasets}
+Dispite the lack of datasets, there are still many datasets listed in section \ref{sec:datasets}
+that can be used to train AST models. Deep learning networks are known to perform better with more
+training data and the greater variety of data will allow the model to generalize better.
+datasets used in the training of our models were TIMIT and CSD. However, there are many genres of
+music that are not represented in children's songs such as rap.
+
 \subsection{Finetuning Wav2Vec 2.0 Feature Extractor}
-\subsection{Adversarial Training}
-\subsection{Transform Singing to Speech Data}
-remove pitch from singing
+The FMA dataset described in section \ref{sec:datasets} can be used to finetune the wav2vec 2.0
+feature extractor. This will allow the model to learn features that are more specific to singing
+and music. This dataset will need some extra preprocessing to detect singing because the songs
+include music with no vocals. This can be done by using a VAD model to detect the singing and
+removing the parts of the audio that do not contain singing. The wav2vec 2.0 paper mentions how
+the model goes through self-supervised learning to learn discrete speech units \citep{wav2vec} and
+a large dataset like FMA will allow the model to learn discrete singing units.
+
+\subsection{Generative Adversarial Networks}
+Generative Adversarial Networks (GANs) are a type of neural network that can be used to generate
+new data and can have recently become popular with the success of Midjourney and
+Stable Diffusion. Voice synthesis and conversion have also improved with services that
+clone voices or text-to-speech services. The intersection of these technologies can
+be used to generate new songs. The songs that are generated do not need to be top highlights
+but plausible sounding songs that can be used to train AST models. Using a GAN can generate
+songs with more variety than the pitch shifted songs mentioned in \ref{sec:pseudoSinging} and
+can lead to better generalization.
+
 \subsection{Phoneme to Grapheme}
-\citep{G2P} or \citep{P2G}
+One of the main applications of AST is to generate and sync lyrics to songs for karaoke
+applications. However, most people cannot read ARPAbet phonemes or IPA phonemes. Instead,
+they can read words and graphemes that form words. Phonemes and graphemes represent the same
+unit of sound but the graphemes are more readable because they are just chunks of the original word.
+This means that as the words are sung, each grapheme of the word can be highlighted individually.
+To get a grapheme output, the phoneme output can be converted to a grapheme output using a
+separate phoneme to grapheme model or generate character level output and align the characters and
+phonemes using the timestamps.
 
